The demographic costs of sexually antagonistic selection in partially selfing populations

### 2021-12-15 doc/SA-Hermaphrodites-wDemography-R1.tex
blueified yesterdays changes
Forgot to put my changes in blue yesterday, sorry, fixed it now

\noindent where $\| \cdot \|$ is the one-norm. The population vector $\tilde{\mbf{n}}(t)$ is projected forward from time $t$ to $t + 1$ by the projection matrix $\tilde{\mbf{A}}(\tilde{\mbf{n}})$ such that 
\begin{linenomath*}
\begin{equation}
	\tilde{\mbf{n}}(t + 1) = \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t)
\end{equation}
\end{linenomath*}
If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix is a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$. In the online supplementary materials we show results for a density-dependent version of the model. In the main text of this paper, we will consider models with density-independent survival and fertility rates. The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.

@@

Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{deVriesCaswell2019a,deVriesCaswell2019b}), see Figure 1 for an example of the possible outcomes in different regions of parameter space. We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection (see left-hand side of Figure \ref{fig:didactic}) and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive (see right-hand side of Figure \ref{fig:didactic}); a situation that we refer to as a 'demographically viable SA polymorphism'.

@@

We used numerical simulation to determine whether a protected SA polymorphism was also demographically viable. Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium. Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model. As a consequence, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}). The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$. We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible \citep{de2020matrix}.

>>>>

\noindent where $\| \cdot \|$ is the one-norm. The population vector $\tilde{\mbf{n}}(t)$ is projected forward from time $t$ to $t + 1$ by the projection matrix $\tilde{\mbf{A}}(\tilde{\mbf{n}})$ such that 
\begin{linenomath*}
\begin{equation}
	\tilde{\mbf{n}}(t + 1) = \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t)
\end{equation}
\end{linenomath*}
\textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix is a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$. In the online supplementary materials we show results for a density-dependent version of the model. In the main text of this paper, we will consider models with density-independent survival and fertility rates. The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.}

@@

Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{deVriesCaswell2019a,deVriesCaswell2019b}), see Figure 1 for an example of the possible outcomes in different regions of parameter space. We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(see left-hand side of Figure \ref{fig:didactic})} and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive \textcolor{blue}{(see right-hand side of Figure \ref{fig:didactic})}; a situation that we refer to as a 'demographically viable SA polymorphism'.

@@

We used numerical simulation to determine whether a protected SA polymorphism was also demographically viable. Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium. \textcolor{blue}{Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model. As a consequence, }the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}). The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$. We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible \citep{de2020matrix}.
